---
title: "CIS 335: Assignment 1"
author: "Stella Sterling"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r load packages, include=FALSE}
#install.packages("tidyverse")
library(tidyverse)
library(kableExtra)
```


## Part 1 & 2

```{r read in data, echo=FALSE}
wineQual <- readr::read_csv(here::here("data","winequalityN.csv"))
```


## Part 3

There are 13 variables: 

| Name    | Attribute   |
|---------|-------------|
| country | nominal   |
| description | nominal |
| designation | nominal |
| points | numeric (ratio) |
| price | numeric (ratio) |
| province | nominal |
| region_1 | nominal |
| region_2 | nominal |
| taster_name | nominal |
| taster_twitter_handle | nominal |
| title | nominal |
| variety | nominal |
| winery | nominal |


## Part 4

```{r mode func, include=FALSE}
#function for mode, is mode not a thing in r??
Mode <- function(x) {
  ux <- unique(x)
  ux[which.max(tabulate(match(x, ux)))]
}
```


Looking at quality points, 'quality': 

```{r sumstat quality, echo=FALSE}
wineQual %>% 
  summarize(
    mean = mean(quality),
    median = median(quality),
    mode = Mode(quality),
    IQR = IQR(quality)) %>% 
  kable(caption="Summary Stats for Quality Points") %>% 
  kable_minimal(full_width=F)
```

```{r points hist, echo=FALSE}
wineQual %>% 
  #filter(!is.na(points)) %>% 
  ggplot(aes(quality)) +
  geom_histogram()
```

Points are looking normal.

```{r part_4 price, echo=FALSE}
wineData %>% 
  filter(!is.na(price)) %>% 
  summarize(
    mean = mean(price),
    median = median(price),
    mode = Mode(price)) %>% 
  kable(caption="Summary Stats for Price") %>% 
  kable_minimal(full_width=F)
```

```{r price hist, echo=FALSE}
wineData %>% 
  filter(!is.na(price)) %>% 
  ggplot(aes(price)) +
  geom_histogram()
```

Ouch. Price is very right skewed. I'm assuming there's a major outlier in the $4000 range. 





